Nanophotonic Inverse Design

### 2011-12-13 presentations/inv_dev_plan_dec-2011/paper.tex
Draft, excluding appendix, completed.

\subsubsection{Low-rank quadratic $f(x)$}
We investigate the case where $f(x) = \|C_1^T x - d_1\|$,
    where $C_1 \in \reals^{n \times p_1}$
    and $d_1 \in \reals^{p_1 \times 1}$, for $p_1 \ll n$.
Here,
    \begin{subequations}\begin{align}
    \nabla f(x) &= C_1 (C_1^T x - d_1) \\
    \nabla^2 f(x) &= C_1 C_1^T
    \end{align}\end{subequations}
    which results in
    \begin{subequations}\begin{align}
    \nabla_x \mathcal{L} &= \rho A^T ((I + \inv{\rho}\tC_1\tC_1^T)Ax
        - b + \inv{\rho}(y - \tC_1 d)) \\
    \nabla_{xx}^2 \mathcal{L} &= \tC_1\tC_1^T + \rho A^T A,
    \end{align}\end{subequations}
    where $\tC_1 = A^{-T}C_1$.

@@

The Newton step is given by
    \begin{equation}
    \Delta x = A^{-1}M ((I + \inv{\rho}\tC_1\tC_1^T)Ax
        - b + \inv{\rho}(y - \tC_1 d)),
    \end{equation}
    and requires $p_1$ solutions of $A^{-T}$ to compute $\tC_1$, 
    and then 1 solution of $A^{-1}$ to compute $\Delta x$.

\subsubsection{Concave, low-rank quadratic $f(x)$}

@@

Newton's method for a linearly-constrained quadratic function is
    \begin{equation}
    \begin{bmatrix} \nabla_{xx}^2 \mathcal{L} & C_2 \\ C_2^T & 0 \end{bmatrix} 
    \begin{bmatrix} \Delta x \\ v \end{bmatrix} = 
    \begin{bmatrix} -\nabla_x\mathcal{L} \\ C_2 x - d_2 \end{bmatrix}.
    \end{equation}
We assume that 
    \begin{subequations}\begin{align}
    (\nabla_{xx}^2 \mathcal{L})^{-1} &= A^{-1} M A^{-T} \\
    \nabla_x \mathcal{L} &= A^T z
    \end{align}\end{subequations}
    and solve for $\Delta x$ 
    using block elimination (see appendix \ref{block elim}),
    \begin{subequations}\begin{align}
    v &= -(\tC_2^T M^{-1} \tC_2)^{-1}(\tC_2^T Ax - d_2 + \tC_2^T Mz) \\
    \Delta x &= -A^{-1} M (z + \tC_2 v),
    \end{align}\end{subequations}
    where $\tC_2 = A^{-T} C_2$.

@@

\subsubsection{Linearly constrained $f(x)$ with arbitrary phase}
\subsubsection{Combined problem}
\subsection{Multi-mode ADMM}
\begin{appendix}
\section{Constructing the relevant matrices and vectors}
\section{Solving the matrices}
\section{Linear algebra definitions}
\subsection{Gradient and Hessian calculations}
\subsection{Matrix inversion lemma}\label{inv lemma}
\begin{equation}
(A + UCV)^{-1} = A^{-1} - A^{-1}U (C^{-1} + V A^{-1} U)^{-1} V A^{-1}
\end{equation}
\subsection{Block elimination of a matrix}\label{block elim}
\begin{equation}
\begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \end{bmatrix} 
\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} = 
\begin{bmatrix} b_1 \\ b_2 \end{bmatrix}
\end{equation}
\begin{subequations}\begin{align}
A_{11} x_1 + A_{12} x_2 &= b_1 \\
x_1 &= A_{11}^{-1} (b_1 - A_{12} x_2)
\end{align}\end{subequations}
\begin{subequations}\begin{align}
A_{21} x_1 + A_{22} x_2 &= b_2 \\
A_{21} A_{11}^{-1} (b_1 - A_{12} x_2) + A_{22} x_2 &= b_2 \\
(A_{22} - A_{21} A_{11}^{-1} A_{12}) x_2 &= b_2 - A_{21} A_{11}^{-1} b_1
\end{align}\end{subequations}
Typically, one computes $A_{11}^{-1} A_{12}$ and $A_{11}^{-1} b_1$, 
    and then solves for the rest of the system.
\end{appendix}
\begin{thebibliography}{99}
\bibitem{ADMM} Boyd group ADMM paper
\end{thebibliography}
\end{document}

>>>>

\subsubsection{Low-rank quadratic $f(x)$}
We investigate the case where $f(x) = \frac{1}{2}\|C_1^T x - d_1\|^2$,
    where $C_1 \in \reals^{n \times p_1}$
    and $d_1 \in \reals^{p_1 \times 1}$, for $p_1 \ll n$.
Here,
    \begin{subequations}\begin{align}
    \nabla f(x) &= C_1 (C_1^T x - d_1) \\
    \nabla^2 f(x) &= C_1 C_1^T
    \end{align}\end{subequations}
    which results in
    \begin{subequations}\begin{align}
    \nabla_x \mathcal{L} &= \rho A^T ((I + \inv{\rho}\tC_1\tC_1^T)Ax
        - b + \inv{\rho}(y - \tC_1 d)) \\
    \nabla_{xx}^2 \mathcal{L} &= \tC_1\tC_1^T + \rho A^T A,
    \end{align}\end{subequations}
    where $\tC_1 = A^{-T}C_1$.

@@

The Newton step is given by
    \begin{equation}
    \Delta x = A^{-1}M ((I + \inv{\rho}\tC_1\tC_1^T)Ax
        - b + \inv{\rho}(y - \tC_1 d_1)),
    \end{equation}
    and requires $p_1$ solutions of $A^{-T}$ to compute $\tC_1$, 
    and then 1 solution of $A^{-1}$ to compute $\Delta x$.

@@

Newton's method for a linearly-constrained quadratic function is
    \begin{equation}
    \begin{bmatrix} \nabla_{xx}^2 \mathcal{L} & C_2 \\ C_2^T & 0 \end{bmatrix} 
    \begin{bmatrix} \Delta x \\ v \end{bmatrix} = 
    \begin{bmatrix} -\nabla_x\mathcal{L} \\ C_2 x - d_2 \end{bmatrix}.
    \end{equation}
We assume that 
    \begin{subequations}\begin{align}
    (\nabla_{xx}^2 \mathcal{L})^{-1} &= A^{-1} M A^{-T} \\
    \nabla_x \mathcal{L} &= A^T z
    \end{align}\end{subequations}
    and solve for $\Delta x$ 
    using block elimination (see appendix \ref{block elim}),
    \begin{subequations}\begin{align}
    v &= -(\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - d_2 + \tC_2^T Mz) \\
    \Delta x &= -A^{-1} M (z + \tC_2 v),
    \end{align}\end{subequations}
    where $\tC_2 = A^{-T} C_2$.
    If we condense the block elimination procedure to a single step,
    we obtain
    \begin{equation}
    \Delta x = -A^{-1} M (z - 
        \tC_2 (\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - d_2 + \tC_2^T Mz)).
    \end{equation}

@@

\subsubsection{Combined performance objective}
As can be seen from the previous examples,
    efficiently computing $\Delta x$ involves transforming $Ax \to z$,
    and then converting back, $A^{-1}z \to x$.
It should be no surprise then, 
    that we can also include all three performance objectives 
    in a single $f(x)$.

This results in computing
    \begin{equation}
    \Delta x = -A^{-1} M (z - 
        \tC_2 (\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - d_2 + \tC_2^T Mz)),
    \end{equation}
    where 
    \begin{subequations}\begin{align}
    M &= I - \tC_1(\rho I + \tC_1^T\tC_1)^{-1}\tC_1^T \\
    \begin{split}
    z &= A^{-T} \nabla_x \mathcal{L} \\
      &= \rho((I + \inv{\rho}\tC_1\tC_1^T)Ax
        - b + \inv{\rho}(y +\tilde{c} - \tC_1 d_1))
    \end{split} \\
    \tilde{c} &= A^{-1} c \\
    \tC_1 &= A^{-1} C_1 \\
    \tC_2 &= A^{-1} C_2. 
    \end{align}\end{subequations}

As expected, the first transformation costs $1 + p_1 + p_2$ solves of $A^{-T}$,
    and the second transformation costs 1 solve of $A^{-1}$.

\subsubsection{Concave, low-rank quadratic $f(x)$}
As a special case, we now consider the choice of 
    $f(x) = - \frac{1}{2}\|C_1^T x \|^2$,
    which is a quadratic, but concave, performance objective.
We consider the case where $f(x)$ is low-rank, that is,
    $C_1 \in \reals^{n \times p_1}$ where $p_1 \ll n$.

To deal with the concavity of $f(x)$ we would like to use
    a truncated Newton method,
    \begin{equation}
    \hat{\mathcal{L}}(x,p,y) = \mathcal{L}(x,p,y) + \frac{\eta}{2} \|x - x_0\|^2
    \end{equation}
    or even
    \begin{equation}
    \Delta x = -(\nabla_{xx}^2 \mathcal{L} + \eta I)^{-1}
        \nabla_x \mathcal{L},
    \end{equation}
    where the coefficient $\eta$ is increased in order to 
    ensure that the augmented Hessian is positive definite 
    (so that $(\nabla_x\mathcal{L})^T\Delta x > 0$).
However, this generally makes inverting the Hessian intractable.

A computationally efficient alternative would be to augment $\mathcal{L}(x,p,y)$
    in the following way,
    \begin{equation}
    \hat{\mathcal{L}}(x,p,y) = \mathcal{L}(x,p,y) + 
        \frac{\eta}{2} \|A(x - x_0)\|^2
    \end{equation}
    so that
    \begin{equation}
    (\nabla_{xx}^2 \hat{\mathcal{L}})^{-1} = 
        A^{-1} (\rho I + \eta I - \tC_1 \tC_1^T)^{-1} A^{-T}
    \end{equation}
    where $\tC_1 = A^{-T} C_1$ and $\nabla_{xx}^2 \mathcal{L} = \rho A^T A$.

For the case $p_1 = 1$, positive definiteness can be ensured by choosing
    $\eta = \| \tC_1 \|^2$.
Generally, for $p_1 > 1$, $\eta$ can be set to the extremal eigenvalue of
    $\tC_1 \tC_1^T$ in order to ensure positive definiteness.
Efficient computation of such an eigenvalue can be performed
    via a $QR$ factorization of $\tC_1$.

The computational cost of using a truncated Newton method is that 
    multiple steps are needed, and so $\Delta x$ must be computed many times.
Fortunately, $\tC = A^{-T} C$ needs only to be computed once,
    bringing the total number of solves to $n_{tn} + p_1$, 
    where $n_{tn}$ is the number of steps needed 
    for the truncated Newton method.

\subsubsection{Linearly constrained $f(x)$ with arbitrary phase}
We now investigate another special case of $f(x)$, 
    namely, when $f(x)$ implements the equality constraint 
    $C_2^T x = e^{i \phi}d_2$,
    where $C_2 \in \reals^{n \times p_2}$ and $p_2 \ll n$.
This situation corresponds to
    the addition of an arbitrary phase $e^{i\phi}$
    to a linear equality constraint.

We simply note that this case can be handled by computing
    \begin{equation}
    \Delta x = -A^{-1} M (z - 
        \tC_2 (\tC_2^T M \tC_2)^{-1}(\tC_2^T Ax - e^{i\phi}d_2 + \tC_2^T Mz)).
    \end{equation}
    for different choices of $\phi$.
The total computational cost is then simply $n_\phi + p_2$ solves,
    where $n_\phi$ is the number of values of $\phi$ which are tried.

\subsection{Multi-mode ADMM}
Since ADMM explicitly separates updating the field and structure variables,
    handling multiple fields is very naturally included in the formulation.
Namely, multiple fields ($x_i$) are updated simultaneously, in parallel,
    and then care ``connected'' by updating a single $p$ 
    which takes all the $x_i$ into account.

\begin{appendix}
\section{Constructing the relevant matrices and vectors}
\section{Solving the matrices}
\section{Linear algebra definitions}
\subsection{Gradient and Hessian calculations}
\subsection{Matrix inversion lemma}\label{inv lemma}
\begin{equation}
(A + UCV)^{-1} = A^{-1} - A^{-1}U (C^{-1} + V A^{-1} U)^{-1} V A^{-1}
\end{equation}
\subsection{Block elimination of a matrix}\label{block elim}
\begin{equation}
\begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \end{bmatrix} 
\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} = 
\begin{bmatrix} b_1 \\ b_2 \end{bmatrix}
\end{equation}
\begin{subequations}\begin{align}
A_{11} x_1 + A_{12} x_2 &= b_1 \\
x_1 &= A_{11}^{-1} (b_1 - A_{12} x_2)
\end{align}\end{subequations}
\begin{subequations}\begin{align}
A_{21} x_1 + A_{22} x_2 &= b_2 \\
A_{21} A_{11}^{-1} (b_1 - A_{12} x_2) + A_{22} x_2 &= b_2 \\
(A_{22} - A_{21} A_{11}^{-1} A_{12}) x_2 &= b_2 - A_{21} A_{11}^{-1} b_1
\end{align}\end{subequations}
Typically, one computes $A_{11}^{-1} A_{12}$ and $A_{11}^{-1} b_1$, 
    and then solves for the rest of the system.
\end{appendix}
\begin{thebibliography}{99}
\bibitem{ADMM} Boyd group ADMM paper
\end{thebibliography}
\end{document}
